---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(readxl)
library(janitor)
library(ggfortify)
library(GGally)
library(modelr)
library(pROC)

```

```{r}

telco_churn <- read_excel("data/telecomms_churn.xlsx") %>% 
  clean_names()

telco_churn

```

```{r}

telco_clean <- telco_churn %>%
  filter(!is.na(total_charges)) %>% 
  select(-customer_id) %>% 
  mutate(across(where(is_character), as_factor)) %>%
  mutate(senior_citizen = as_factor(if_else(senior_citizen == 1, "Yes", "No")),
         churn = if_else(churn == "Yes", TRUE, FALSE))

```

```{r, message=FALSE, warning=FALSE}

split1 <- telco_clean %>%
  select(gender, senior_citizen, partner, churn)

split2 <- telco_clean %>%
  select(dependents, tenure, phone_service, churn)

split3 <- telco_clean %>%
  select(internet_service, contract, monthly_charges, total_charges, churn)


split1 %>% 
  ggpairs(progress = FALSE)

split2 %>% 
  ggpairs(progress = FALSE)

split3 %>% 
  ggpairs(progress = FALSE)

```

Model 1: churn ~ senior_citizen
Model 2: churn ~ tenure
Model 3: churn ~ total_charges

```{r}

model1_glm <- glm(churn ~ tenure, 
                  data = telco_clean,
                  family = binomial(link = "logit")
                  )

model2_glm <- glm(churn ~ senior_citizen, 
                  data = telco_clean,
                  family = binomial(link = "logit")
                  )

model3_glm <- glm(churn ~ total_charges, 
                  data = telco_clean,
                  family = binomial(link = "logit")
                  )


```

```{r}

summary(model1_glm)
summary(model2_glm)
summary(model3_glm)

```

```{r}

model1_with_pred <- telco_clean %>%
  select(tenure, churn) %>% 
  add_predictions(model1_glm, type = "response")

roc_obj_1 <- model1_with_pred %>% 
  roc(response = churn, predictor = pred)



model2_with_pred <- telco_clean %>%
  select(senior_citizen, churn) %>% 
  add_predictions(model2_glm, type = "response")

roc_obj_2 <- model2_with_pred %>% 
  roc(response = churn, predictor = pred)


model3_with_pred <- telco_clean %>%
  select(total_charges, churn) %>% 
  add_predictions(model3_glm, type = "response")

roc_obj_3 <- model3_with_pred %>% 
  roc(response = churn, predictor = pred)

ggroc(data = list("tenure_model" = roc_obj_1,
                  "senior_citizen_model" = roc_obj_2,
                  "total_charges_model" = roc_obj_3),
      legacy.axes = TRUE) +
    labs(x = "False Positive Rate",
       y = "True Positive Rate")

```

```{r}

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```

Model 1 (churn ~ tenure) has the highest AUC therefore is most likely to be the best classifier for determining the churn rate of a customer.

```{r}

summary(model1_glm)

```

The intercept (b0) is 0.03 which suggests the sigmoid function is slightly to the left. The coefficient of x is minus which suggests that as tenure increases the probability of churning decreases.

```{r}

min_tenure <- min(telco_clean$tenure)
max_tenure <- max(telco_clean$tenure)

log_predictions <- tibble(tenure = seq(min_tenure, max_tenure, 1)) %>% 
  add_predictions(model1_glm,
                  type = "response")

ggplot(telco_clean) +
  geom_jitter(aes(x = tenure, y = as.numeric(churn)), alpha = 0.5,
              height = 0.1, width = 0.1) +
  geom_line(data = log_predictions, aes(x = tenure, y = pred), col = "red") +
              labs(y = "Estimated p(churn)")

```

Odd ratio

```{r}

odds_change <- function(b1, change){
  
  exp(b1 * change)
  
}

odds_change(-0.039010, 1)

```

For every increase in tenure by 1 month, the odds of churning is 4% lower than before.