---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(gtools)

titanic <- read_csv("data/titanic.csv")

```
# MVP 
## 1.1 The Titanic


```{r}

titanic <- titanic %>% 
  mutate(is_crew = if_else(str_detect(class, "[0-9]"), FALSE, TRUE))

```
Calculate the conditional probabilities prob(survived∣crew) an
prob(survived∣passenger) to settle this dispute.

```{r}

prob_crew <- titanic %>% 
  filter(is_crew) %>% 
  summarise(prob = n()/nrow(titanic))

prob_passenger <- titanic %>% 
  filter(!is_crew) %>% 
  summarise(prob = n()/nrow(titanic))

prob_survived_and_crew <- titanic %>% 
  filter(is_crew & survived == "yes") %>% 
  summarise(prob = n()/nrow(titanic))

prob_survived_and_passenger <- titanic %>% 
  filter(!is_crew & survived == "yes") %>% 
  summarise(prob = n()/nrow(titanic))

frac_crew_survived <- prob_survived_and_crew / prob_crew
frac_passenger_survived <- prob_survived_and_passenger / prob_passenger

frac_crew_survived # 23.7%
frac_passenger_survived # 38.0

```

## 1.2 Probability Consolidation Questions


```{r}

source("prob.R")

```


### Question 1

If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}

one_roll <- rolldie(1)

prob_lt_5 <- one_roll %>% 
  filter(X1 < 5) %>% 
  summarise(prob = n() / nrow(one_roll))

prob_lt_5

```

### Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
#first method, use rolldie() function
num_outcomes <- nrow(rolldie(3))
#second method, use permutations() from gtools
num_outcomes_verify <- nrow(permutations(6,3, repeats.allowed = TRUE))

num_outcomes
num_outcomes_verify
```

### Question 3.
In an experiment where we roll four dice:

- how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?
- what is the probability of this event?

```{r}
# all possible permutations of 4 die rolls
four_rolls <- rolldie(4)

# filter on values of 4 rolls equals 5, 6 or 7.
four_rolls %>% 
  mutate(sum = rowSums(.)) %>% 
  filter(sum %in% c(5,6,7)) %>% # 34 outcomes
  summarise(prob = n() / nrow(four_rolls))
  
```

- what assumptions do we apply in answering part ii?

We have assumed that we are using fair dice (i.e. each number is equally likely)
and that we are using the total number of permutations rather than the number of
combinations fo dice rolls (i.e. we are rolling them one-by-one rather than all
at once!)

### Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}

num_permutations <- nrow(permutations(5,5))
#only one correct order therefore number of correct solutions is 1, probability is 1 over the number of permutations
prob_desc_height <- 1 / num_permutations
prob_desc_height

#this can be verified using an example, we are interested in the order of {5,4,3,2,1} representing descending height order.
perm <- as.tibble(permutations(5,5))

prob_desc_height_2 <- perm %>% 
  filter(V1 == 5,
         V2 == 4,
         V3 == 3,
         V4 == 2,
         V5 == 1) %>% 
  summarise(prob = n() / nrow(perm))

prob_desc_height_2

```

### Question 5.
I roll six dice:

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

```{r}

num_permutations <- nrow(permutations(6, 6, repeats.allowed = TRUE))
#only one correct order therefore number of correct solutions is 1, probability is 1 over the number of permutations
prob_desc_num <- 1 / num_permutations
prob_desc_num



```
what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

```{r}

num_123456 <- nrow(permutations(6,6, repeats.allowed = FALSE))

prob_1_to_6 <- num_123456 / num_permutations

prob_1_to_6

```

```{r}
# row must contain 6, 5,..., 1, but can occur in any order
contains_654321 <- six_rolls %>%
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = FALSE))

prob <- nrow(contains_654321) / nrow(six_rolls)
prob

```


### Question 6.
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

```{r}

playing_cards <- cards() %>% 
  mutate(colour = if_else(suit %in% c("Spade", "Club"), "black", "red"))

prob_black_and_spades <- playing_cards %>%
  filter(colour == "black" & suit == "Spade") %>% 
  summarise(prob = n() / nrow(playing_cards))

prob_black <- playing_cards %>% 
  filter(colour == "black") %>% 
  summarise(prob = n() / nrow(playing_cards))

prob_black_and_spades <- prob_black_and_spades / prob_black

prob_black_and_spades #probability is 50% that the black card is spades

```


```{r}

?choose(1, 6)

```

