---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(infer)

msleep

```

# MVP

## 1.1 Hypothesis testing - practical

### Question 1
Explore the dataset and familiarise yourself with it.

```{r}

glimpse(msleep)
summary(msleep)

```

### Question 2
Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate statistical test to determine whether the mean sleep_total in the sampled population of animal types differs from the typical value for jabberwockies.

Determine the hypotheses we want to test:
$$
H_0: \mu_{sleep\ jabberwockies} = \mu_{sleep\ of\ all\ animals}\\
H_1: \mu_{sleep\ jabberwockies} \neq \mu_{sleep\ of\ all\ animals}
$$

This test will be a two-tailed test.

Set alpha = 0.05

Firstly, identify the point estimate (sample statistic) that we want to compare,
i.e. mean of sleep_total, in this case it is provided to us as 7 hours.

```{r}
# calculate sample statistic, i.e. mean of total sleep for all animals in sample
sleep_observed_stat <- msleep %>% 
  
  summarise(mean_sleep_total = mean(sleep_total))


```


Create null distribution from the sample data using the bootstrap method and 
number of reps = 10,000. We will be using the "infer" package workflow to do 
this.
```{r}

null_distribution_sleep <- msleep %>% 
  specify(response = sleep_total) %>% 
  hypothesise(null = "point", mu = 7) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")

# create histogram of null_distribution
null_distribution_sleep %>% 
  visualise() +
  shade_p_value(obs_stat = sleep_observed_stat, direction = "both")

```

```{r}

p_value <- null_distribution_sleep %>% 
  get_p_value(obs_stat = sleep_observed_stat, direction = "both")
  
```

The p-value is much less (p < 0.001) than the alpha value (0.05) therefore we will 
_reject_ the null hypothesis. This indicates that the average total_sleep is 
_significantly_ different from 7 hours (the average total sleep for 
jabberwockies).

### Question 3
Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

Determine the hypotheses we want to test:
$$
H_0: \mu_{sleep\ omnivores} = \mu_{sleep\ herbivores}\\
H_1: \mu_{sleep\ omnivores} > \mu_{sleep\ herbivores}
$$
Can also be written as:
$$
H_0: \mu_{sleep\ omnivores} - \mu_{sleep\ herbivores} = 0\\
H_1: \mu_{sleep\ omnivores} - \mu_{sleep\ herbivores} > 0
$$

This test will be a one-tailed test.

Set alpha = 0.05

Firstly, identify the point estimate (sample statistic) that we want to compare,
i.e. mean of omnivores total sleep minus the mean of herbivores total sleep.

```{r}
# filter sample data for only omnivore and herbivore animals
msleep_filtered <- msleep %>% 
  filter(vore %in% c("omni", "herbi"))

# calculate sample statistic, i.e. mean of total sleep for all animals in sample
vore_observed_stat <- msleep_filtered %>% 
  specify(response = sleep_total, explanatory = vore) %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

vore_observed_stat

```


Create null distribution from the sample data using the "permute" method and 
number of reps = 10,000.

```{r}

null_distribution_vore <- msleep_filtered %>% 
  specify(response = sleep_total, explanatory = vore) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 10000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

# create histogram of null_distribution
null_distribution_vore %>% 
  visualise() +
  shade_p_value(obs_stat = vore_observed_stat, direction = "right")

```

```{r}

p_value <- null_distribution_vore %>% 
  get_p_value(obs_stat = vore_observed_stat, direction = "right")
  
p_value

```
The p-value (p = 0.1256) is greater than the alpha value (0.05) therefore we 
_fail to reject_ the null hypothesis. This indicates that the difference between
omnivore and herbivore average total sleep is _not statistically significant_.

### Question 4
Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than 5%.

Determine the hypotheses we want to test:
$$
H_0: \pi_{domesticated} = 5\%\\
H_1: \pi_{domesticated} > 5\%
$$

This test will be a one-tailed test.

Set alpha = 0.05

Firstly, identify the point estimate (sample statistic) that we want to compare,
i.e. mean of omnivores total sleep minus the mean of herbivores total sleep.

```{r}
# create column specifying if animal is domesticated, change NA to FALSE
msleep_dom <- msleep %>%
  mutate(is_domesticated = conservation == "domesticated",
         is_domesticated = coalesce(is_domesticated, FALSE),
         .after = conservation)

# calculate sample statistic, i.e. proportion of domesticated animals
dom_observed_stat <- msleep_dom %>%
  specify(response = is_domesticated, success = "TRUE") %>% 
  calculate(stat = "prop")

dom_observed_stat

```


Create null distribution from the sample data using the "permute" method and 
number of reps = 10,000.

```{r}

null_distribution_dom <- msleep_dom %>% 
  specify(response = is_domesticated, success = "TRUE") %>% 
  hypothesise(null = "point", p = 0.05) %>% 
  generate(reps = 10000, type = "draw") %>% 
  calculate(stat = "prop")

# create histogram of null_distribution
null_distribution_dom %>% 
  visualise() +
  shade_p_value(obs_stat = dom_observed_stat, direction = "right")

```

```{r}

p_value <- null_distribution_dom %>% 
  get_p_value(obs_stat = dom_observed_stat, direction = "right")
  
p_value

```
The p-value (p = 0.0095) is less than the alpha value (0.05) therefore we 
_reject_ the null hypothesis. This indicates that there is evidence to support
the alternative hypothesis that the average proportion of domesticated animals 
is greater than 5%.

## 1.2 Hypothesis testing - Defining the Hypothesis

For the following three business problems write out:

- What kind of test you would use
- H0 and Ha in both mathematical notation and in words.
- Also state the method you would use to generate the null distribution (bootstrap, permutation or draw).

### Question 1

Test type: One-sample type - proportions - one-tailed

Hypotheses: 

Null hypothesis: The number of people in the town who are aware of the coffee 
shop is less than 40%.

The alternative hypothesis: The number of people in the town who are aware of
the coffee shop is greater than or equal to 40%

$$
H_0 : \pi_{aware} < 40\% \\
H_1 : \pi_{aware} \geq 40\% 
$$

Null distribution method: "draw"

### Question 2

Test type: Two sample test - difference in proportions

Hypotheses: 

Null hypothesis: The CTR of Group A and Group B are the same, i.e. the location
of the banner on the webpage shows no change in the CTR.

The alternative hypothesis: The CTR of Group A and Group B are not the same, 
i.e. the location of the banner on the webpage does show a change in the CTR.

$$
H_0 : \pi_{CTR\ A} = \pi_{CTR\ B} \\
H_1 : \pi_{CTR\ A} < \pi_{CTR\ B} 
$$

Null distribution method: "permute"

### Question 3

Test type: one sample test - two-tailed - mean

Hypotheses: 

Null Hypothesis: Car part width is equal to 145mm

Alternative Hypothesis: Car part width is not equal to 145mm

$$
H_0 : \mu_{width} = 145mm \\
H_1 : \mu_{width} \neq 145mm 
$$

Null distribution method: "bootstrap"

## 1.3 Hypothesis Testing - Interpreting the results

### Question 1.

Coffee shop problem. Significance level: 0.05, calculated p-value: 0.07

Since the p-value is greater than the significance we should 
_reject the alternative hypothesis_ which states that the average proportion of 
people in the town who are aware of the coffee shop is greater than or equal to 
40% and _fail to reject_ the null hypothesis, i.e. less than 40% of people are
aware of the coffee shop.

### Question 2.

Website company problem. Significance level: 0.01, p-value: 0.006

Since the p-value is less than the significance level we must _reject_ the null
hypothesis since we have evidence to support the alternative hypothesis. In this
case we have evidence to support that the change of banner location has had a 
statistically significant impact on the CTR.

### Question 3.

Manufacturing company problem. Significance level: 0.05, p-value: 0.55

Since the p-value is greater than the significance level we should _fail to reject_ 
the null hypothesis which states that the manufacturing company is achieving the
quality assurance target of 145mm for that car part, i.e. there is no statistical
significant deviation from the 145mm average target.

# EXTENSION

## 2.1 Market Basket Analysis

```{r}
library(janitor)

transactions <- read_csv("data/online_retail_subset.csv") %>% 
  clean_names()

```

For the first section we are interested in the purchase of two particular items:

item A - ‘HEART OF WICKER SMALL’ (StockCode 22469)
item B - ‘LARGE CAKE TOWEL PINK SPOTS’ (StockCode 21110)


### Question 1

Calculate the support for item A (this will be the support for a single item)

```{r}
# filter data for stock code and distinct invoice_no
num_transactions_A <- transactions %>% 
  filter(stock_code == 22469) %>% 
  distinct(invoice_no, stock_code) %>% 
  summarise(prob_A = n())

# calculate number of total distinct invoices
total_transactions <- transactions %>% 
  distinct(invoice_no) %>% 
  summarise(n())

sup_A <- num_transactions_A / total_transactions

sup_A

```

### Question 2

Calculate the support and confidence for rule (A→B).

```{r}
# find the number of invoices that have both products A and B
num_A_and_B <- transactions %>% 
  filter(stock_code %in% c(22469, 21110)) %>% 
  group_by(invoice_no) %>% 
  summarise(num_prod = n_distinct(stock_code)) %>% 
  filter(num_prod == 2) %>% 
  summarise(num_A_and_B = n())

# calculate sup(A -> B)
sup_A_and_B <- num_A_and_B / total_transactions
  
sup_A_and_B

  # filter(invoice_no %in% c(536520, 536592, 537144, 537773, 538071, 538349, 538524),
  #        stock_code %in% c(22469, 21110))

```

```{r}
# conf(A -> B) = sup(A -> B) / sup(A)
conf_A_and_B <- sup_A_and_B / sup_A
conf_A_and_B

```

### Question 3.
Calculate the lift for (A → B)

```{r}

num_transactions_B <- transactions %>% 
  filter(stock_code == 21110 ) %>% 
  distinct(invoice_no, stock_code) %>% 
  summarise(prob_B = n())

sup_B <- num_transactions_B / total_transactions

sup_B

```

```{r}

lift_A_and_B <- sup_A_and_B / (sup_A * sup_B)

lift_A_and_B

```

# Apriori Algorithm

```{r}

library(arules)
library(arulesViz)
transactions_reformat <- transactions %>%
  select(invoice_no, description) %>%
  na.omit()

write_csv(transactions_reformat, "transactions_reformat.csv")

apriori_format <- read.transactions("transactions_reformat.csv", format = "single", sep = ",", header = TRUE, cols = c("invoice_no", "description"))

```

```{r}
inspect(head(apriori_format))
```

```{r}
transactions %>% 
  distinct(description)

```


```{r}
# get the rules

rules <- apriori(apriori_format, parameter = list(supp = 0.02, conf = 0.6, 
                                                  minlen = 3),
                 control = list(verbose = F))

# rules <- sort(rules, by="confidence", decreasing=TRUE)

inspect(rules[])

```

```{r}

library(arulesViz)

plot(rules, method="graph", engine = "interactive")

```

