---
title: "R Notebook"
output: html_notebook
---

# Question 1
Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}

library(tidyverse)
library(CodeClanData)

backpack

```

# Question 2
What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}

backpack %>% 
  ggplot(aes(backpack_weight)) +
  geom_bar() +
  labs(x = "Backpack Weight (lb)",
       y = "Count",
       title = "Frequency of backpack weights amongst students\n") +
  scale_y_continuous(labels = scales::label_number(accuracy = 1, scale = 1),
                     breaks = seq(0, 12), #12 is manually entered here, ideally this would be calculated.
                     expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(0, max(backpack$backpack_weight), by = 5)) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"))

```
The frequency plot (histogram) of backpack weights amongst students shows a 
relatively normal distribution with the majority of weights between 4 and 21lbs.
There are however outliers in the dataset with a maximum weight of 35lbs.

# Question 3
Are there differences in backpack weight between sex?

```{r}

backpack %>% 
  ggplot(aes(x = backpack_weight,
             y = sex,
             fill = sex)) + 
  geom_boxplot(show.legend = FALSE) +
  labs(x = "Backpack Weight (lb)",
       title = "Spread of backpack weights between genders\n") +
  scale_x_continuous(breaks = seq(5, 35, 5)) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        axis.title.x = element_blank(),
        axis.text.x = element_text(face = "bold"),
        panel.grid.major.y = element_line(colour = "gray90")) +
  coord_flip()

```
The boxplots above show that although the average of backpack weight is similar
between the sexes, approx. 10lbs for females and 11lbs for males. However, the
difference between the genders arises from the range of weight values observed. 
For males the range of backpack weights is 3 to 35lbs and for females it is
3 - 25lbs.

# Question 4.

What is the relationship between body weight and backpack weight?

```{r}

backpack %>% 
  ggplot(aes(x = body_weight,
             y = backpack_weight)) +
  geom_point(alpha = 0.4) +
  labs(x = "Body Weight (lb)",
       y = "Backpack Weight (lb)",
       title = "Relationship between body and backpack weight") +
  theme(panel.background = element_rect(fill = "white"),
        panel.border = element_rect(colour = "gray", fill = NA),
        axis.line = element_line(colour = "gray90"),
        panel.grid = element_line(colour = "gray90")) +
  geom_smooth(n = 2, se = F, colour = "black")

```
The scatterplot above shows that there appears to be somewhat of a positive
correlation towards the centre of the data where higher backpack weights are
observed with persons of higher body weight. This, in principal, would make
sense since heavier people are likely to be stronger and can carry heavier 
backpacks and vice versa for lighter people. However, across the entire dataset
this relationship falls apart due to the outliers. The line of best fit shows 
that there is no correlation between the two parameters.

# Question 5.
Is this relationship different between men and women? Choose an appropriate plot
and interpret your findings.

```{r}


backpack %>% 
  ggplot(aes(x = body_weight,
             y = backpack_weight,
             group = sex,
             colour = sex)) +
  geom_point(show.legend = F, alpha = 0.4) +
  labs(x = "Body Weight (lb)",
       y = "Backpack Weight (lb)",
       title = "Relationship between body and backpack weight",
       subtitle = "Split by gender") +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major.y = element_line(colour = "gray90")) +
  facet_wrap(~ sex, ncol = 2) +
  geom_smooth(n = 2, se = F, show.legend = F)

```
Splitting the data between Male and Female shows that there is a small positive 
relationship between body and backpack weight, this is different from the
analysis done previously which indicated no correlation. For males, there is 
still no correlation between the parameters.

# Question 6.
What proportion of students report back pain?

```{r}
backpack %>%
  group_by(back_problems) %>% 
  summarise(count = n())

```


```{r}

backpack %>% 
  group_by(back_problems) %>% 
  summarise(percent = n()/nrow(backpack)) %>% 
  ggplot(aes(x = "", y = percent, label = paste0(percent*100, "%"))) +
  geom_col(position = "fill",
           fill = c("gray","red")) +
  coord_polar(theta = "y") +
  geom_text(size = 5, position = position_fill(vjust = 0.5)) +
  labs(title = "Proportion of students reporting back problems",
       fill = "Back Problems") +
  theme(panel.background = element_rect(fill = "white"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank())
  
```
The pie chart above shows that 32% of students experience back problems.


# Question 7.

Is this proportion different between men and women?

```{r}

backpack %>% 
  group_by(sex, back_problems) %>% 
  summarise(count = n()) %>% 
  group_by(sex) %>% 
  summarise(count, sum = sum(count), percent = count/sum)

```

```{r}
backpack %>%
  group_by(sex, back_problems) %>% 
  summarise(count = n()) %>% 
  group_by(sex) %>% 
  summarise(count, sum = sum(count), frac = count/sum * 100)


backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>% 
  ggplot(aes(x = sex,
              group = back_problems,
              fill = back_problems,
             label = mean(body_weight))) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "gray90")) +
  scale_y_continuous(labels = scales::percent,
                     expand = c(0, 0)) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major.y = element_line(colour = "gray90"),
        axis.title = element_blank(),
        axis.text.x = element_text(face = "bold")) +
  labs(title = "Proportion of students suffering from back problems",
       subtitle = "Split by gender",
       fill = "Back Problems?")
  

```
The graph above shows that females experience a higher proportion of back
problems (44%) vs males (18%).

[Unable to get the percentages on the graph so added in the y-axis to show
proportion. Table above shows percentage values for each gender.]

# Question 8.

Now you’ve investigated your data separately, create plots to describe the
relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}

backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>% 
  ggplot(aes(y = backpack_weight, x = body_weight, group = back_problems)) +
  geom_point(aes(colour = sex, shape = back_problems),
             size = 2,
             fill = "white") +
  scale_shape_manual(values = c(16, 21)) +
  # facet_grid(sex ~ back_problems, labeller = label_both) +
  labs(y = "Backpack Weight (lb)",
       x = "Body Weight (lb)",
       fill = "Gender",
       title = "Relationship between body and backpack weights",
       subtitle = "Split by gender and reported back problems",
       shape = "Back Problems?",
       colour = "Gender") +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major = element_line(colour = "gray90"),
        axis.title = element_text(face = "bold"),
        legend.key = element_blank())

```
The above plot consolidates all of the data related to body, backpack, gender 
and back problems into a single plot. The plot follows the same trend as in Q4 
and Q5. The rate at which back problems occurs at different levels between 
males and females, females typically report back problems at lower backpack 
weights (bottom left of the plot), which is consistent with lower typical body weights too.

average ratio of back problems between males and females
```{r}

backpack %>%
  mutate(back_problems = as.logical(back_problems)) %>% 
  group_by(sex, back_problems) %>% 
  summarise(avg_ratio = mean(ratio)) %>% 
  ggplot(aes(x = sex, y = avg_ratio, fill = sex)) +
  geom_col() +
  facet_wrap(~ back_problems, labeller = labeller(back_problems = c("TRUE" = "Back Problems: TRUE", "FALSE" = "Back Problems: FALSE"))) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major = element_line(colour = "gray90"),
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "Gender", 
       y = "Average Bag to Body Weight Ratio\n",
       fill = "Gender",
       title = "Comparison of average bag to body weight ratios",
       subtitle = "And the impact on back problems") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 0.1))

```
The bar chart above shows that back problems tend to occur at higher bag to body weight ratio. This makes sense that if a student is carrying a heavy bag (in relation to their body weight) then there is a higher risk of back problems.

For females, the bag to body weight ratio is higher in instances where back 
problems are reported than for males. The data suggests that females on average carry heavier bags (in relation to their body weight, i.e. ratio) than males do.
This is because, on average, females are lighter than males and are quite likely
to carry the same weight of bag.

### EXTENSION

# Question 1.

Does reported backpain differ across the year groups? Ignore year 0 in your data.

```{r}

backpack %>% 
  filter(back_problems == 1) %>% 
  ggplot(aes(x = year, y = back_problems)) +
  geom_col() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,13),
                     breaks = scales::pretty_breaks()) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major.y = element_line(colour = "gray90"),
        legend.key = element_blank()) +
  labs(x = "Year group", 
       y = "Reports of backpain (No.)",
       title = "Reports of Backpain across each year group")
```
The data shows that there is quite a significant increase in reports of backpain
between year groups 1 and 2, this may correlate to the number of things required
for classes in 3rd year. However, this data doesn't account for the total number
of students in each dataset, i.e. doesn't indicate the proportion of students in
each year group, jsut the number of students reporting backpain.



# Question 2.

Is there a difference between how heavy the bags those in different years carry?

```{r}

backpack %>% 
  group_by(year) %>% 
  summarise(avg_backpack_weight = mean(backpack_weight, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = avg_backpack_weight)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = c(0,0),
                     breaks = scales::pretty_breaks()) +
  scale_x_continuous(breaks = 0:6) +
  theme(panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "gray"),
        panel.grid.major.y = element_line(colour = "gray90"),
        legend.key = element_blank()) +
  labs(x = "Year group", 
       y = "Average backpack weight (lb)",
       title = "Average backpack weight across each year group")


```
The average backpack weight increases significantly from year groups 0 to 1 and
plateaus at Year group 2. Year groups 2 - 4 are very similar (approx. 12lb) with
a further increase in Year Group 5 up to 13lb, before a significant drop down to around 8lb for Year Group 6. The data shows a good correlation with what might 
be expected, i.e. first year of college has the least amount of materials to be 
carried, years 2-4 are very similar and year 5 (final year) is the most 
demanding therefore needs the most materials (i.e. heavier backpack).
