---
title: "R Notebook"
output: html_notebook
---

# MVP
Using the dataset austen_books() from the package janeaustenr.

```{r}

library(tidyverse)
library(tidytext)
library(textdata)
library(janeaustenr)

```

```{r}

titles <- c("Pride & Prejudice", "Sense & Sensibility") 
books <- list(prideprejudice, sensesensibility)

books <- purrr::map_chr(books, paste, collapse = " ")

```

```{r}

two_books <- tibble(
  title = titles,
  text = books
) %>% 
  unnest_tokens(word, text)

two_books

```

## Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}

two_books %>% 
  count(title, word) %>% 
  group_by(title) %>% 
  slice_max(n, n = 3)

```

The most common words across both books are "the", "to" and "of.

## Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.

```{r}

two_books %>% 
  anti_join(stop_words) %>% 
  count(title, word) %>% 
  group_by(title) %>% 
  slice_max(n, n = 3)

```

The most common words (excluding stop words) are "Elizabeth", "Darcy" and "Bennet" in Pride& Prejudice, and "Elinor", "Marianne" and "Time" in Sense & Sensibility.

## Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r}
# using bing list as sentiment words
two_books %>% 
  inner_join(lexicon_bing()) %>% 
  group_by(title, word) %>% 
  summarise(word_count = n()) %>% 
  slice_max(word_count, n = 3)

```
The most common sentiment words from P&P are "miss", "well" and "good" and for S&S are "well", "miss" and "good". 

# Extension
Taking your results above. Can you create a plot which visualises the differences between the books?

```{r}

pride_book <- tibble(
  id = 1:length(prideprejudice),
  lines = prideprejudice
) %>% 
  unnest_tokens(word, lines)

pride_book_sentiments <- pride_book %>% 
  inner_join(lexicon_afinn()) %>% 
  group_by(id) %>% 
  summarise(mean_sentiment = mean(value)) %>% 
  mutate(id = id/ max(id),
         title = "Pride & Prejudice")


sense_book <- tibble(
  id = 1:length(sensesensibility),
  lines = sensesensibility
) %>% 
  unnest_tokens(word, lines)

sense_book_sentiments <- sense_book %>% 
  inner_join(lexicon_afinn()) %>% 
  group_by(id) %>% 
  summarise(mean_sentiment = mean(value)) %>% 
  mutate(id = id/ max(id),
         title = "Sense & Sensibility")

combined_sentiments <- pride_book_sentiments %>% 
  bind_rows(sense_book_sentiments)

combined_sentiments %>% 
  ggplot(aes(x = id, y = mean_sentiment, colour = title)) +
  geom_smooth(se = FALSE) +
  labs(x = "Percentage of story complete (%)",
       y = "Average sentence sentiment score",
       title = "How does sentiment change through the books?",
       col = "Book Title") +
  scale_x_continuous(labels = scales::percent, expand = c(0,0)) +
  scale_y_continuous(expand = c(0.2, 0)) +
  theme_classic()

```

S&S starts off more positive than P&P but sharply decreases before rising again at 50% of the book. It rises a bit then decreases again before finally rising towards the end of the book. P&P drops more slowly as the book progresses with a low-point at around 70% of the story completed, before rising again continuously until the end of the book.

Interestingly, both books start happier than they end (in terms of sentimental words used!)