---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(janitor)

```
# Question 1.
Load the data, clean_names() and explore it.
```{r}

mobile_ratings <- read_csv("data/20190928-items.csv") %>% 
  clean_names()

mobile_ratings
```

# Question 2.
Find the brand with the highest numbers of phones in the dataset.

```{r}

mobile_ratings %>% 
  group_by(brand) %>% 
  summarise(num_phones = n()) %>% 
  slice_max(num_phones)

```
Samsung has the highest number of phones in the dataset

# Question 3.
For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}

samsung_ratings <- mobile_ratings %>% 
  filter(brand == "Samsung") %>% 
  dplyr::select(rating)

samsung_rating_stats <- samsung_ratings %>% 
  summarise(mean_rating = mean(rating),
            sd_rating = sd(rating))

samsung_ratings %>% 
  ggplot(aes(rating)) +
  geom_histogram(aes(y = ..density..), col = "white") +
  stat_function(
    fun = ~dnorm(., mean = samsung_rating_stats$mean_rating,
                 sd = samsung_rating_stats$sd_rating),
    col = "red"
  )

```
The normal distribution provides a reasonable estimate of the data, there is a concentration of values near the mean although there are some imperfections in the fit, specifically the two tall peaks near the mean and the "outlier" rating values. Overall, this data tends to follow a normal distribution.


```{r}

library(MASS)

fitdistr(samsung_ratings$rating, densfun = "normal")

```
The sd value is slightly different from the output of the sd() function used above.

```{r}

samsung_ratings_z_scored <- samsung_ratings %>% 
  mutate(z = (rating - samsung_rating_stats$mean_rating) / samsung_rating_stats$sd_rating)

# number of ratings within one sd
sr_1sd <- samsung_ratings_z_scored %>% 
  filter(abs(z) <= 1) %>% 
  summarise(frac = n() / nrow(samsung_ratings_z_scored)) %>%
  pull()
# 79.3%

# number of ratings within two sd
sr_2sd <- samsung_ratings_z_scored %>% 
  filter(abs(z) <= 2) %>% 
  summarise(frac = n() / nrow(samsung_ratings_z_scored)) %>%
  pull()
# 91.9%

# number of ratings within two sd
sr_3sd <- samsung_ratings_z_scored %>% 
  filter(abs(z) <= 3) %>% 
  summarise(frac = n() / nrow(samsung_ratings_z_scored)) %>%
  pull()
# 97.9%


```

```{r}

nd_1sd <- pnorm(q = 1) - pnorm(q = -1)
nd_2sd <- pnorm(q = 2) - pnorm(q = -2)
nd_3sd <- pnorm(q = 2) - pnorm(q = -3)
```

```{r}

as_tibble(
  list(
  sd = c(1,2,3),
  samsung_ratings_data = c(sr_1sd, sr_2sd, sr_3sd),
  normal_dist = c(nd_1sd, nd_2sd, nd_3sd))
)

```
The samsung ratings data doesn't appear to match the proportions of data 
contained within the 1, 2 and 3 standard deviations. The samsung data shows a 
much higher proportion of the data being captured within one sd suggesting that
the data is quite heavily concentrated near the mean in comparison to the normal
distribution, this combined with a meaningful amount of data points at the 
extremes (i.e. rating of 1) is enough to effect the mean and standard deviation
and the amount of data contained within the 1 standard deviation range.

However, at 2 sds the amount of data is less than what a normal distribution 
would expect. Again, this is presumably related to the number of "outliers" in 
the samsung dataset. For 3sds there is close agreement between the two values, 
where the majority of the data points are included within the +/-3sd range.

# EXTENSION
Investigate the use of the qqnorm() plot in R to determine whether data is normally distributed. What is a ‘quantile-quantile plot’? What do we expect to see in a qqnorm() plot if data is nearly normally distributed?

```{r}

qqnorm(samsung_ratings$rating)
qqline(samsung_ratings$rating)

```
The QQ-plot above shows that there is significant deviation from the Samsung
rating data and normal distribution. If the the data was perfectly normal then
the data (shown as circles) would follow the normal distribution line. Whilst 
the data appears to follow this line in the middle of the data set, there are 
significant differences towards the extremes, this is evident at the lower
sample ratings (e.g. 1 and 2 ratings). Therefore, the QQ-plot would suggest that
the data does not follow a normal distribution due to the disproportionate 1 and
2 score ratings in the dataset, also known as having "heavy tails".

```{r}

library(car)
qqPlot(samsung_ratings$rating)

```

