---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(stringr)

code_clan_tweets <- read_csv("data/code_clan_tweets.csv")

```

# Question 1

```{r}

glimpse(code_clan_tweets)

```
# Question 2

Find total number of favourites, omit quoted tweets.

```{r}

code_clan_tweets %>% 
  filter(is_quote == F) %>% 
  summarise(total_favourites = sum(favorite_count))
  
```

# Question 3

Summarise mean number of retweets, split by platform type, omitting quoted
tweets.

```{r}

code_clan_tweets %>%
  filter(is_quote == F) %>% 
  group_by(source) %>% 
  summarise(avg_retweet_count = sum(retweet_count))

```

# Question 4

Count total number of likes, split by media type and arrange in descending
order, assume tweets without listed media type are "text".

```{r}

code_clan_tweets %>% 
  mutate(media_type = if_else(is.na(media_type), "text", media_type)) %>% 
  group_by(media_type) %>% 
  summarise(sum_favs = sum(favorite_count)) %>% 
  arrange(desc(sum_favs))

```

# Question 5

Find mean number of character in CodeClan tweet.

```{r}

code_clan_tweets %>% 
  summarise(avg_tweet_length_text = mean(nchar(text)),
            avg_tweet_length_display_width = mean(display_text_width))


```

The two methods do not give the same results, this is because the "display_text_
width" column sums the characters which are not hashtags or urls.

# Question 6

```{r}

code_clan_info <- read_csv("data/code_clan_info.csv")

```


```{r}

code_clan_tweets_joined <- code_clan_tweets %>% 
  left_join(code_clan_info, by = "tweet_id")

view(code_clan_tweets_joined)

```

# Question 7

```{r}

codeclan_hashtags <- code_clan_tweets_joined %>% 
  select(tweet_id, hashtags) %>% 
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  filter(!is.na(hashtags))

codeclan_hashtags

```

### Extension

# Question 8

```{r}
view(codeclan_hashtags)
```


```{r}
codeclan_hashtags %>% 
  mutate(hashtags_subbed = str_sub(hashtags, start = 1, end = 2)) %>% 
  filter(str_detect(hashtags_subbed,"c\\("))

```

# Question 9

```{r}

code_clan_tweets %>% 
  filter(str_detect(text, "(?i)Edinburgh")) %>% 
  summarise(count = n())

```

```{r}
code_clan_tweets %>% 
  filter(str_detect(text, "Edinburgh|edinburgh")) %>% 
  summarise(count = n())


```

# Question 10

number of "@"s

```{r}

pattern <- "@[A-Za-z0-9_]+"

code_clan_tweets %>% 
  select(text) %>%
  mutate(text = str_to_lower(text)) %>% 
  str_extract_all(pattern) %>% 
  flatten_chr() %>% 
  tibble() %>% 
  mutate(cleaned_name = str_sub(.,2,-1)) %>% 
  group_by(cleaned_name) %>% 
  summarise(num_tags = length(cleaned_name))

```

