---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)

```
# Question 1
```{r}

housing_prices <- read_csv("data/housing_prices.csv")

housing_prices

```

# Question 2

```{r, warning=FALSE}

library(GGally)

ggpairs(housing_prices)

```

# Question 3 

```{r}

housing_prices <- housing_prices %>% 
  select(-total_bedrooms)

```
# Question 4

## i

Done above

## ii
```{r}

housing_prices %>% 
  summarise(cor = cor(median_income, median_house_value))

# strong correlation between median_income and median_house_value = 0.642

housing_prices %>% 
  ggplot(aes(x= median_income, y = median_house_value)) +
  geom_point()

```

# Question 5

```{r}

housing_prices %>% 
  distinct(ocean_proximity)

```

5 distinct ocean_proximity categories therefore we would expect n - 1 dummy variables, i.e. 4 dummy variables for ocean_proximity from this dataset.

# Question 6

```{r}

model <- lm(median_house_value ~ median_income,
            housing_prices)

```

```{r}
library(ggfortify)

autoplot(model)

```

Plot one shows good randomness between residuals and fitted values
Plot two shows that the residuals don't quite fit a perfect normal distribution, i.e. deviance from the straight line at higher theoretical quantiles
Plot three shows consistent standardised residuals across the dataset - perfect!

The plots show that the line of best fit from the model is acceptable.

# Question 7

Add proximity to the beach (categoric variable to the analysis)

```{r}

library(fastDummies)

housing_prices %>% 
  dummy_cols(select_columns = "ocean_proximity",
             remove_first_dummy = TRUE,
             remove_selected_columns = TRUE)

```

```{r}

model_income_rooms <- lm(median_house_value ~ median_income + total_rooms,
                       housing_prices)

autoplot(model_income_rooms)

summary(model_income_rooms)

```

```{r}


model_income_oceanprox <- lm(median_house_value ~ median_income + ocean_proximity,
                housing_prices)

summary(model_income_oceanprox)

```

```{r}

autoplot(model_income_oceanprox)

```

Residual values look ok however the residuals still don't match perfectly with a normal distribution.

From the summary table there are positive increases in the median house price when the property is located near a bay or ocean with a much larger increase if the property is located on an island, conversely inland properties tend to have lower median house values. 

The change in median_house_values can be explained by around 55% from the median_income and ocean_proximity values.

# Question 8


```{r}

model_log_inc <- lm(median_house_value ~ median_income + ocean_proximity +
                      log(median_income):ocean_proximity,
                    housing_prices)

summary(model_log_inc)

```

I don't believe that the interaction between log(median_income) and ocean_proximity is valid since it is hard to believe that these two variables would interact. This is further emphasised by the negligible increase in the R-squared value from 0.5529 (without interaction) and 0.5578 (with interaction).

The issue may lie with the use of the log being applied to the income value, the income values already appear to be in some transformed state(?).

# Question 8

```{r}

autoplot(model_log_inc)

```

```{r}

library(mosaic)

plotModel(model_log_inc)

```

Only the Island data points adopt a logarithmic curve-like shape when the interactions are included in the model, the remaining are (close to) parallel lines.

