---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(janitor)

```

```{r}

national_incidence <- read_csv("raw_data/incidence_at_scotland_level.csv") %>% 
  clean_names()

glimpse(national_incidence)

```

```{r}

hb_names <- read_csv("raw_data/geography_codes_and_labels_hb2014_01042019.csv") %>% 
  clean_names() %>% 
  select(hb, hb_name)

hb_incidence <- read_csv("raw_data/incidence_by_health_board.csv") %>% 
  clean_names() %>% 
  left_join(hb_names, "hb") %>% 
  filter(hb_name == "NHS Borders") %>% 
  select(-ends_with("_qf"))



hb_incidence
hb_names

hb_incidence <- hb_incidence %>% 
  left_join(hb_names, "hb") %>% 
  filter(hb_name == "NHS Borders") %>% 
  select(-ends_with("_qf"))

```


# Total annual cancer incidences in the borders

```{r}

hb_incidence %>%
  filter(cancer_site == "All cancer types" & sex == "All") %>%
  ggplot(aes(x = year, y = incidences_all_ages)) +
  geom_line() +
  theme_classic() +
  scale_y_continuous(limits = c(0,NA)) +
  labs(x = "Year",
       y = "Total Incidences",
       title = "Total number of cancer incidences in the Scottish Borders",
       subtitle = "1996 to 2020")

```
The year on year incidence rates of cancer has trended upwards since 1996 reaching
a peak in 2017, since 2017, however, there has been a significant downwards trend
with the incidences in 2020 being the lowest since 2009, albeit this is presumably due 
to the covid pandemic where people were advised not to go to hospitals except in 
emergencies and many cancers may have gone un-diagnosed during this time period.



# All cancer rates compared with europe and world

```{r}

hb_incidence %>%
  filter(cancer_site == "All cancer types" & sex == "All") %>%
  select(id:year, crude_rate, easr, wasr) %>% 
  pivot_longer(c(crude_rate, easr, wasr), names_to = "type", values_to = "incidence_rate") %>% 
  ggplot(aes(x = year, y = incidence_rate, group = type, col = type)) +
  geom_line()


```

```{r}

hb_incidence %>%
  filter(cancer_site == "All cancer types" & sex == "All") %>% 
  filter(sir_lower95pc_confidence_interval > 100 |
           sir_upper95pc_confidence_interval < 100)


```

```{r}

X5yr_summary_incidence_by_cancer_network_region <- read_csv("raw_data/5yr _summary_incidence_by_cancer_network_region.csv") %>% 
  clean_names()

X5yr_summary_incidence_by_cancer_network_region %>% 
  distinct(region)

```

```{r}

hb_5yr <- read_csv("raw_data/5yr_summary_incidence_by_health_board.csv") %>% 
  clean_names() %>% 
  left_join(hb_names, "hb") %>% 
  select(id, hb_name, everything(), -hb, -ends_with("_qf")) %>% 
  filter(hb_name == "NHS Borders")

top_5_cancers_sex <- hb_5yr %>% 
  filter(cancer_site != "All cancer types", sex == "All") %>% 
  slice_max(incidences_all_ages, n = 5) %>% 
  mutate(cancer_site = factor(cancer_site))

top_5_cancers_list <- top_5_cancers_sex %>% 
  pull(cancer_site)

top_5_cancers_sex %>% 
  ggplot(aes(x = reorder(cancer_site, sort(incidences_all_ages)),
             y = incidences_all_ages, label = scales::comma(incidences_all_ages))) +
  geom_col() +
  geom_text(position = position_nudge(y = 75)) +
  theme_classic() +
  scale_y_continuous(limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1)),
                     labels = scales::comma) +
  labs(x = "Cancer Type",
       y = "Incidences",
       title = "Top 5 most common cancers in the Scottish Borders",
       subtitle = "2016 - 2020") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

```

```{r}

hb_5yr %>% 
  arrange(desc(sir_lower95pc_confidence_interval))

```
age groups of total cancer incidences
```{r}
top_5_list <- top_5_cancers_sex %>% 
  pull(cancer_site)


hb_5yr_ages <- hb_5yr %>% 
  filter(sex != "All") %>% 
  mutate(sex = factor(sex, c("Male", "Female"))) %>% 
  select(id:incidences_age85and_over) %>% 
  pivot_longer(incidences_age_under5:incidences_age85and_over,
               names_to = "age_group", values_to = "incidences") %>% 
  mutate(age_group = str_remove(age_group, "incidences_age|"),
         age_group = str_remove(age_group, "_"),
         age_group = str_replace(age_group, "to", " - "),
         age_group = str_replace(age_group, "under5", "Under 5"),
         age_group = str_replace(age_group, "85andover", "85+")) 

age_group_list <- hb_5yr_ages %>% 
  head(18) %>% 
  pull(age_group)

hb_5yr_ages %>% 
  mutate(age_group = factor(age_group, levels = age_group_list)) %>% 
  ggplot(aes(x = age_group, y = incidences)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0), labels = scales::comma) +
  theme_classic()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Age Group",
       y = "Total Incidences",
       title = "Total number of cancer incidences in the Scottish Borders",
       subtitle = "1996 to 2020")
```



```{r}
hb_5yr_incidence_rates <- hb_5yr %>% 
  filter(sex != "All") %>% 
  mutate(sex = factor(sex, c("Male", "Female"))) %>% 
  pivot_longer(incidence_rate_age_under5:incidence_rate_age85and_over,
               names_to = "age_group", values_to = "incidences") %>% 
  select(id:sex, age_group, incidences) %>% 
  mutate(age_group = str_remove(age_group, "incidence_rate_age|"),
         age_group = str_remove(age_group, "_"),
         age_group = str_replace(age_group, "to", " - "),
         age_group = str_replace(age_group, "under5", "Under 5"),
         age_group = str_replace(age_group, "85andover", "85+")) 

hb_5yr_incidence_rates %>% 
  mutate(age_group = factor(age_group, levels = age_group_list)) %>% 
  ggplot(aes(x = age_group, y = incidences)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0), labels = scales::comma) +
  theme_classic()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Age Group",
       y = "Incidence Rate (per x,000)",
       title = "Cancer incidence rates per age group in the Scottish Borders",
       subtitle = "1996 to 2020")

```

sex of top 5 cancers 
```{r}

hb_5yr %>% 
  filter(cancer_site %in% top_5_list, sex != "All") %>% 
  select(id:sex, incidences_all_ages) %>% 
  mutate(sex = factor(sex, c("Male", "Female")),
         cancer_site = factor(cancer_site, levels = rev(top_5_cancers_list))) %>% 
  ggplot(aes(x = cancer_site, y = incidences_all_ages, fill = sex)) +
  geom_col(position = "fill", col = "black") +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  labs(y = "Proportion (%)",
       title = "Proportion of top 5 cancer types by gender in Scottish Borders",
       subtitle = "Incidences recorded between 2016 and 2020",
       fill = "Gender") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(face = "bold")) +
  coord_flip()

```


# Final graph - which cancers to pay close attention too in future
i.e. cancers which are higher than the national average and affect many people are priority

```{r}
library(plotly)

```


```{r}

p <- hb_5yr %>% 
  select(cancer_site, sex, incidences_all_ages, 
         standardised_incidence_ratio:sir_upper95pc_confidence_interval) %>% 
  filter(sex == "All", incidences_all_ages > 0) %>% 
  mutate(standardised_incidence_ratio = round(standardised_incidence_ratio, 1)) %>% 
  rename(SIR = standardised_incidence_ratio,
         Incidences = incidences_all_ages,
         Type = cancer_site) %>% 
  ggplot(aes(label = Type, y = Incidences, x = SIR)) +
  geom_point() +
  xlim(0,200) +
  ylim(0, 10000) +
  scale_y_continuous(trans = "log10", name = "log(incidences)") +
  theme_classic() +
  geom_vline(aes(xintercept = 100)) +
  geom_hline(aes(yintercept = 100)) +
  theme(panel.background = element_rect(colour = "black")) +
  labs(x = "SIR",
       y = "Incidences",
       title = "Cancer type prioritisation matrix",
       subtitle = "NHS Borders data: 2016 - 2020")



ggplotly(p,
         tooltip = c("label", "y", "x")) %>%
  config(displayModeBar = F)

```

```{r}
hb_5yr %>% 
  select(cancer_site, sex, incidences_all_ages, crude_rate,
         standardised_incidence_ratio) %>% 
  arrange(standardised_incidence_ratio)
```

