---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(CodeClanData)
library(janitor)
library(scales)

```

```{r}

qb_revenue_breakdown
qb_monthly_sales
qb_competitors
qb_device_data

```

# 2 MVP

# 2.1 Part 1

# Question 1.
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}

qb_revenue_breakdown %>%
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             fill = product)) +
  geom_bar(stat = "identity") +
  labs(x = "\nYear",
       y = "Revenue",
       fill = "Product",
       title = "Composition of Annual Revenue 2018 & 2019\n")

```

# Question 2.
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}

qb_monthly_sales %>% 
  clean_names() %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, 
            y = amount)) +
  geom_line() +
  labs(x = "Date",
       y = "Sales Amount",
       title = "Monthly sales amount in the last year")

```

# Question 3.
Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}

qb_monthly_sales %>% 
  clean_names() %>%
  filter(cashflow != "Kwikbit Unit 1 Sales") %>%
  ggplot(aes(x = date,
             y = amount,
             colour = cashflow)) +
  geom_line() +
  labs(x = "Date",
       y = "Amount",
       title = "Revenue and costs in the last year")
  
  
```

# Question 4.
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}

qb_competitors %>% 
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             fill = company)) +
  geom_area() +
  labs(x = "Year",
       y = "Revenue",
       title = "Competitor Revenues - last 5 years",
       fill = "Company")


```

# Question 5.
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}

qb_competitors %>% 
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             colour = company)) +
  geom_line() +
  geom_point() +
  labs(x = "Year",
       y = "Revenue",
       title = "Competitor Revenues - last 5 years",
       colour = "Company")


```

# Question 6.
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}

qb_device_data %>% 
  mutate(time = (hours * 60 + mins)/60) %>% 
  ggplot(aes(x = time,
         y = counts,
         colour = id)) +
  geom_line() +
  facet_grid(id ~ device) +
  labs(x = "Time",
       y = "Steps",
       colour = "ID")

```

# 2.2 Part 2

# Question 1.
Take the plots that you produced in part one and now polish them by:

Adding appropriate labels
Changing the scales and coordinates when appropriate.

Applying a unified theme, which is described below:
Graphs should have white backgrounds, and use colour sparingly.
There should be faint grid lines.
Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
All plots should use colours from the following company colour scheme.
```{r}

source("hw_theme.R")
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

```

# Q1.1

```{r}

qb_revenue_breakdown %>%
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             fill = product)) +
  geom_bar(stat = "identity") +
  labs(x = "\nYear",
       fill = "Product",
       title = "Composition of Annual Revenue 2018 & 2019\n") +
  scale_y_continuous("Revenue (£)\n",
                     labels = label_comma()) +
  scale_fill_manual(
      values = col_scheme) +
  hw_theme()

```

#Q2.1

```{r}

qb_monthly_sales %>% 
  clean_names() %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, 
            y = amount)) +
  geom_line(colour = "#BD7040") +
  labs(x = "\nDate",
       y = "Sales Amount (£)\n",
       title = "Monthly sales amount in the last year\n") +
  scale_colour_manual(values = col_scheme) + # this line doesn't work
  hw_theme()

```

#Q3.1

```{r}

qb_monthly_sales %>% 
  clean_names() %>%
  filter(cashflow != "Kwikbit Unit 1 Sales") %>%
  ggplot(aes(x = date,
             y = amount,
             colour = cashflow)) +
  geom_line() +
  labs(x = "Date",
       title = "Revenue and costs in the last year",
       colour = "Type") +
  scale_y_continuous("Amount (£)\n",
                     labels = label_comma()) +
  scale_colour_manual(values = col_scheme) +
  hw_theme()


```

#Q4.1 

```{r}

qb_competitors %>% 
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             fill = company)) +
  geom_area() +
  labs(x = "\nYear",
       y = "Revenue",
       title = "Competitor Revenues - last 5 years\n",
       fill = "Company") +
  scale_y_continuous("Revenue (£)\n",
                     labels = label_comma()) +
  scale_fill_manual(values = col_scheme) +
  hw_theme()


```

#Q5.1 

```{r}

qb_competitors %>% 
  clean_names() %>% 
  ggplot(aes(x = year,
             y = revenue,
             colour = company)) +
  geom_line() +
  geom_point() +
  labs(x = "",
       y = "Revenue (£)\n",
       title = "Competitor Revenues - last 5 years\n",
       colour = "Company") +
  scale_colour_manual(values = col_scheme) +
  scale_y_continuous(labels = label_comma()) +
  hw_theme()


```

#Q6.1 

```{r}

qb_device_data %>% 
  mutate(time = (hours * 60 + mins)/60) %>% 
  ggplot(aes(x = time,
         y = counts,
         colour = id)) +
  geom_line() +
  facet_grid(id ~ device) +
  labs(x = "Time",
       y = "Steps",
       colour = "ID",
       title = "Daily Steps pattern per person and device") +
  scale_colour_manual(values = col_scheme) +
  hw_theme()

```

